Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-11-09 paper/introduction.tex
updatess to introduction

@@

Consider the example shown in Fig.~\ref{fig:motivating-eg}.
\textbf{TODO: Replace by more complex example where data-structure is updated
over multiple invocations of method and finally transferred.}
This code represents the logic for a streaming query operator (also known
as a continuous query). The operator receives a continuous stream of
input messages, which are batched into time-windows associated with a logical
timestamp. Messages may be delivered out-of-order and multiple messages with
the same timestamp may be delivered. A timing-message, periodically indicates that
no more input messages with a timestamp less than a given value $t$ will be subsequently
delivered. At this point, the operator can create the output message corresponding
to time $t$ and deliver it to its successors.

@@

The goal of this work is an automatic (memory-safe) region-based memory management technique
that supports dynamic regions and higher-order functions.

>>>>

\newcommand{\TODO}[1]{\textbf{TODO: #1}}

@@

Consider the example shown in Fig.~\ref{fig:motivating-eg}.
\TODO{Replace by more complex example where data-structure is updated
over multiple invocations of method and finally transferred.}
This code represents the logic for a streaming query operator (also known
as a continuous query). The operator receives a continuous stream of
input messages, which are batched into time-windows associated with a logical
timestamp. Messages may be delivered out-of-order and multiple messages with
the same timestamp may be delivered. A timing-message, periodically indicates that
no more input messages with a timestamp less than a given value $t$ will be subsequently
delivered. At this point, the operator can create the output message corresponding
to time $t$ and deliver it to its successors.

@@

The goal of this work is an automatic (memory-safe) region-based memory management technique
that supports dynamic regions and higher-order functions.
Though our focus is on memory-safety, dynamic regions also have other side-benefits.
When a producer transfers a data-structure to a consumer, in a shared-memory context , our approach
provides an ownership discipline that ensures that the producer cannot subsequently modify
the data-structure. When the same transfer happens in a distributed setting, the system guarantees
that the data-structure is self-contained and can be safely transmitted across address spaces.
\TODO{How accurate is this claim?}

The key memory safety property we wish to ensure is that there are no dangling references:
i.e., a reference to an object that has been freed.
\TODO{Or, should we say: no dangling reference is dereferenced}
A key invariant and restriction that we utilize to achieve this is that no object $o_1$ in one dynamic
region $D_1$ can contain a pointer to an object $o_2$ in another dynamic region $D_2$.
(In cases where this is too restrictive, the user has to resort to (deep) copying $o_2$ to $D_1$.
This is conceptually quite reasonable given the perspective that $D_1$ is a self-contained data-structure.
Furthermore, this is quite analogous, from a performance perspective, to the copying that happens
when a garbage collector promotes an object from a lower generation to a higher generation.
From this perspective, a region may be seen as a user-controlled generation.)
